---
title: "Methods used to set management objectives for nutrients - ECOSTAT Questionnaire"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    highlight: kate
  pdf_document:
    template: my_html_document (template = "/Library/Frameworks/R.framework/Versions/4.0/Resources/library/rmdformats/templates/template.html")
    toc: yes
    toc_depth: 3
    keep_tex: yes
---

```{r setup, include=FALSE}
library(rmdformats)
library(tinytex)
library(rmarkdown)
#library(pandoc)

library(tidyverse)
library(tidytext)
library(here)
library(knitr)
library(kableExtra)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

# custom format function which first calls rmarkdown::html_document with the "default" template (to retain MathJax functionality), but then modifies the pandoc template setting
my_html_document <- function(template = "path_to_custom_template", ...) {
  base_format <- rmarkdown::html_document(...)

  template_arg <- which(base_format$pandoc$args == "--template") + 1L
  base_format$pandoc$args[template_arg] <- template

  base_format
}

# Get output document type and use as the format argument in kable function
table_format = if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
```

ECOSTAT has a mandate to compare methods used by Member States to set management objectives  for nutrients that are compatible with good ecological status and good environmental status.

This questionnaire aims to understand national approaches for setting and implementing management objectives for nutrients and to help Member States and collaborating countries to improve their approaches and, thereby, achieve the objectives of the Water Framework Directive (WFD) and the Marine Strategy Framework Directive (MSFD).

```{r load data, include=FALSE}
FName.dat <-here::here("Data","PartA.csv")
FName.datB <-here::here("Data","PartB.csv")
FName.datC <-here::here("Data","PartC.csv")

dat.PartA<- read.csv(file = FName.dat, sep=";", header = TRUE) #fileEncoding="latin1"
dat.PartB<- read.csv(file = FName.datB, sep=";", header = TRUE) #fileEncoding="latin1"
dat.PartC<- read.csv(file = FName.datC, sep=";", header = TRUE) #fileEncoding="latin1"
```

```{r pre-treatment A, include=FALSE}
dat.PartA <- type.convert(dat.PartA) #convert data automatic as appropriate

dat.PartA$BE<-as.integer(dat.PartA$BE)
dat.PartA$BE.Br<-as.integer(dat.PartA$BE.Br)
dat.PartA$BE.Fl<-as.integer(dat.PartA$BE.Fl)

dat.PartA$WaterCategory<-factor(dat.PartA$WaterCategory, levels=c("Lakes","Rivers","Transitional","Coastal","Marine")) #order factor levels

PartA.long <-pivot_longer(dat.PartA, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.long <- PartA.long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

PartA.long$Acronym <- as.factor(PartA.long$Acronym)
```

#### Overview of countries participationg in questionnaire

Twenty-six countries participated in this survey, five of which are land-locked countries reporting only for freshwater categories.

```{r countries data, }
dat.countries <- read.csv(file = here::here("Data","Countries.csv"), sep=";", header = TRUE)
dat.countries$Type <- factor(dat.countries$Type, levels = c("EU Member States", "EFTA Member States", "EU Candidate Countries"))

#left join countries full name and info
PartA.long <- left_join( PartA.long, dat.countries, by = "Acronym", keep = FALSE) 
PartA.long$Countries <- as.factor(PartA.long$Countries)

Tab1 <- PartA.long %>% filter(Reply == 1)%>%
  select(c(WaterCategory, Acronym, Countries))%>%
  count(WaterCategory, Acronym, Countries)%>%
  mutate(n = ifelse(n < 100, "y", n)) #replace number of replies just to get countries per wat cat

#Table 1  
options(knitr.kable.NA = '')
Tab1 %>% spread(key = WaterCategory, value = n) %>%
  knitr::kable(format = table_format,col.names = c("Acronym","  ","Lakes","Rivers","Transitional","Coastal","Marine"), align = "llccccc", caption = "Table 1. Countries* participating in survey per water category.") %>%#,table.attr = "style='width:70%;'"
  kableExtra::kable_styling (position = "c") %>%
  kableExtra::add_header_above(header = c("Country" = 2, "Water Category" = 5),align = "centre")

```

 *Iceland and Norway (EFTA Members) and Turkey (EU candidate) are not EU Member States. 

## How are management objectives for nutrients derived (methods)? (Part A)

## Q1. Have you set management objectives for nutrients for waters bodies that are exposed to nutrient pollution?

```{r Q1,fig.cap="Figure 1. Number of countries responding, overview by water category."}
#select data Q1
PartA.Q1 <- dat.PartA %>% filter(QuestionNo_b==1)

PartA.Q1$ResponseOptionText <- str_trim(PartA.Q1$ResponseOptionText)

#run function to display the sample size at y= -1
give.n <- function(x){
  return(c(y = -1, label = x))#length(x)
}
  
plotQ1 <- ggplot(PartA.Q1, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotQ1
```

```{r Countries detail Q1, include=FALSE}
PartA.Q1long <-pivot_longer(PartA.Q1, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.Q1long <- PartA.Q1long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.count1 <- PartA.Q1long %>% 
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

#function to extract countries list per reply
CntryListfunction <- function(dat) {
      #i is a unique WaterCategory - ResponseOptionText combination
      #dat is a subset of the data of interest
  WCreply_levels<- unique(dat$WCreply)
  Lists<-list()# prepare a container
       #loop through WaterCategory - ResponseOptionText unique combinations
  for (i in 1:length(WCreply_levels)) {
  acronym_list<- dat$Acronym[which(dat$WCreply==WCreply_levels[i])]
  Lists[i] <- list(acronym_list)
  #print(Lists)
  }
  return(Lists)
}

unique(id.count1$WCreply) #levels with country replies
xx1 <-CntryListfunction (id.count1)#countries replying
#xx?[[11]] # use [[?]] to extract values directly in text
```

Most countries have set management objectives for all water bodies exposed to nutrient pollution in the different water categories, with few countries considering only some of the nutrient polluted water bodies (`r xx1[[2]]` in lakes; `r xx1[[5]]` rivers; `r xx1[[8]]` in transitional waters; and `r xx1[[11]]` in marine waters). Only `r xx1[[3]]` in lakes, `r xx1[[6]]` in rivers, and `r xx1[[12]]` in marine waters have not set management objectives for water bodies exposed to nutrient pollution. Results show that France (FR) does not follow a standardised national approach, varying according to water agency (but confirm this).

(tbc Martyn) Add here a critical analysis paragraph for freshwater categories.
I don't think that there is anything that needs to be added here.   
<<<<<<< HEAD
=======
Try again
>>>>>>> 2195281edf58a864e90d28003c6538451fe2c835

(tbc Heliana) Add here a critical analysis paragraph for marine water categories.
 
## Q2. When setting management objectives for nutrients, is there a focus on reduction of phosphorus, nitrogen, or both or on the limiting nutrient?

```{r Q2, fig.cap="Figure 2. Number of countries responding, overview by water category."}
#select data Q2
PartA.Q2 <- dat.PartA %>% filter(QuestionNo_b==2)

PartA.Q2$ResponseOptionText <- str_trim(PartA.Q2$ResponseOptionText)
  
plotQ2 <- ggplot(PartA.Q2, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotQ2
```

```{r Countries detail Q2, include=FALSE}
PartA.Q2long <-pivot_longer(PartA.Q2, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.Q2long <- PartA.Q2long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.count2 <- PartA.Q2long %>%
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

unique(id.count2$WCreply) #levels with country replies
xx2 <-CntryListfunction (id.count2)#countries replying
#xx?[[11]] # use [[?]] to extract values directly in text
```

Responses reflect, to a large extent, an interplay between biogeography and local pressures. Most countries focus on reducing both nitrogen and phosphorus whatever the water category considered; those who focussed on one or the other did so because local circumstances meant that this was the most likely reason for problems. That would be, for example, the case of phosphorus reduction in lakes, adopted by nine countries (`r xx2[[1]]`), or nitrogen in lakes (`r xx2[[2]]`) and rivers (`r xx2[[6]]`) or in coastal (`r xx2[[13]]`) and marine waters (`r xx2[[16]]`). UWWTD, ND and WFD were all cited as regulatory drivers for these actions. Responses focussed on the consequences of nutrients on transitional water ecosystems, but there was also some acknowledgment of the role of transitional waters as conduits of nutrients to coastal and marine waters too.

The limiting nutrient approach is more commonly assumed in freshwater categories (lakes: `r xx2[[4]]`; rivers: `r xx2[[8]]`) than in marine influenced systems (transitional: `r xx2[[12]]`, coastal: `r  xx2[[15]]`).  What is not clear from the response is how the limiting nutrient is determined in any given situation.  Several responses referred to prevailing dogma that nitrogen and phosphorus are the nutrients most likely to be limiting (and, therefore, responsible for eutrophication when present in excess) and some (e.g. PL) indicated that targets were to achieve concentrations consistent with the good/moderate boundary.  This, in turn, presumes a good quantitative understanding of causal relationships between nutrients and measures of the condition of the flora and fauna which can be difficult to obtain for transitional waters (Salas-Herrero et al., 2019).   It can probably be assumed that exceeding supporting element standards for nitrogen and phosphorus is the default criterion for recognising a need for nutrient management.  However, as Phillips et al. (2018) point out, there are a variety of approaches to setting nutrient standards, with different levels of precaution.   No mention is made of more targeted, case specific approaches in the responses; however, this does not mean that they are not used in individual countries.

## Q3. If you have set management objectives for nutrients, please indicate which methodology has been used. 

```{r Q3, fig.cap="Figure 3. Number of countries responding, overview by water category."}
#select data Q3
PartA.Q3 <- dat.PartA %>% filter(QuestionNo_b==3)

PartA.Q3$ResponseOptionText <- str_trim(PartA.Q3$ResponseOptionText)
  
plotQ3 <- ggplot(PartA.Q3, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotQ3
```

```{r Countries detail Q3, include=FALSE}
PartA.Q3long <-pivot_longer(PartA.Q3, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.Q3long <- PartA.Q3long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.count3 <- PartA.Q3long %>%
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

unique(id.count3$WCreply) #levels with country replies
xx3 <-CntryListfunction (id.count3)#countries replying
#xx?[[11]] # use to extract values directly in text
```

The most commonly used method to derive nutrient targets in all water categories was based on modeled pressure-response relationships between nutrient(s) and biological quality elements (BQE) (add formula to id countries in each Water category here?). According to their description, phytoplankton (chlorophyll) and macrophytes (confirm this) were the BQEs that countries have most frequently (?) considered for establishing such pressure-response relationships. 

Add more details on the type of modeling, rational, BQEs selected, and modeled variables (loads?, etc..)
Freshwaters: the four possible answers are not exclusive, with several countries using more than one approach.  For example, a modelling approach might be combined with historical information by an expert to give a more nuanced target than would be provided by any approach in isolation.  The term “modelling”, itself, covers a wide range of possibiltilies.  BQE-response relationships are often strong in lakes and are well-suited to deriving standards following the methods in the Best Practice Guide [how should this be cited here?] but a range of other modelling approaches are employed.   Few respondents mentioned which BQE was used.   Of those who did, phytoplankton, phytobenthos, macrophytes, invertebrates and fish were all mentioned, depending on specific contexts.  There is also a distinction between those countries that set targets in terms of concentrations and those who use loads.  A further distinction is between those who used mass balance models and those who used dynamic process models although details in questionnaire responses did not always explain the mechanisms behind their models.   Lakes are well suited to the use of historical information, through the use of palaeoecology; historical information is also used in rivers by eight countries, though none of these provided more information on how this was done. 

(tbc Martyn/Heliana) Historical data and expert judgment are also frequently used methods...

(tbc Martyn/Heliana) Add here a critical analysis paragraph.

Countries have provided a short description of the methodology, including links to background reports or publications that can be consulted for further details in the report compiling "EC ECOSTAT Nutrients Questionnaire Responses" (add link to report). 

## Q4. Do you determine the maximum critical load or a load reduction target that is compatible with good status for nutrient sensitive BQEs? 

```{r Q4, fig.cap="Figure 4. Number of countries responding, overview by water category."}
PartA.Q4 <- dat.PartA %>% filter(QuestionNo_b==4)

PartA.Q4$ResponseOptionText <- str_trim(PartA.Q4$ResponseOptionText)

plotQ4 <- ggplot(PartA.Q4, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col(aes(fill=ResponseOptionText)) +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotQ4 + theme(legend.position = "none") 
```

```{r Countries detail Q4, include=FALSE}
PartA.Q4long <-pivot_longer(PartA.Q4, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.Q4long <- PartA.Q4long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.count4 <- PartA.Q4long %>%
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

unique(id.count4$WCreply) #levels with country replies
xx4 <-CntryListfunction (id.count4)#countries replying
#xx?[[11]] # use to extract values directly in text
```

Nutrient critical loads or load reduction targets compatible with good status of nutrient sensitive BQEs are only set by half or less of the reporting countries in lakes (`r xx4[[1]]`), rivers (`r xx4[[3]]`), transitional (`r xx4[[5]]`), coastal (`r xx4[[7]]`) and marine waters (`r xx4[[9]]`).

(tbc Martyn/Heliana)  Based on open replies, add here a paragraph on What methodology do you use?
(tbc Martyn/Heliana)  Add here a critical analysis paragraph (see the open replies).

Detailed responses to this question were ambiguous, with 16 (lakes) and 18 (rivers) answering “no” to the question, implying that they did not determine the maximum critical load or a load reduction target, whereas some of the more comments send by these same countries suggested that they did.  What seems to be clear is that many see target setting as a process that needs to be applied to individual water bodies, rather than derived as a guide value for all water bodies within a broadly-defined type.   Where targets are defined as concentrations rather than loads there is an inference that some form of mass balance model is also employed in order to work out how a target may be achieved for any individual water body.

## Q5. Is there a relationship between the management objectives set for nutrients and the nutrient boundaries set for ecological status? 

```{r Q5, fig.cap="Figure 5. Number of countries responding, overview by water category."}
#select data Q5
PartA.Q5 <- dat.PartA %>% filter(QuestionNo_b==5)

PartA.Q5$ResponseOptionText <- str_trim(PartA.Q5$ResponseOptionText)

PartA.Q5$ResponseOptionText <- case_when(PartA.Q5$ResponseOptionText == "Yes, the management objective is to reach the nutrient concentration at good/moderate boundary" ~ "Yes, management objective is to reach the nutrient concentration at G/M boundary",
                                        PartA.Q5$ResponseOptionText == "Yes, the management objective is to reach the mid-point of good status (between the high/good and the good/moderate boundary) as defined by the biological quality elements" ~ "Yes,  management objective is to reach the mid-point of G status as defined by the BQEs",
                                      PartA.Q5$ResponseOptionText == "Yes, the management target is to reach the mid-point of good status (between the high/good and the good/moderate boundary) as defined by the biological quality elements" ~ "Yes,  management objective is to reach the mid-point of G status as defined by the BQEs",
                                      TRUE ~ as.character(PartA.Q5$ResponseOptionText))

plotQ5 <- ggplot(PartA.Q5, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotQ5
```

```{r Countries detail Q5, include=FALSE}
PartA.Q5long <-pivot_longer(PartA.Q5, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.Q5long <- PartA.Q5long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.count5 <- PartA.Q5long %>%
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

unique(id.count5$WCreply) #levels with country replies
xx5 <-CntryListfunction (id.count5)#countries replying
#xx?[[11]] # use to extract values directly in text
```

Most countries acknowledge that there is a relationship between the management objectives set for nutrients and the nutrient boundaries set for ecological status, and most commonly the management objective is to reach the nutrient concentration at good/moderate boundary. In some cases the management objective is  slightly more ambitious and aims to reach the mid-point between the high/good and the good/moderate boundary as defined by the biological quality elements (lakes: `r xx5[[2]]`, rivers:  `r xx5[[5]]`, transitional:  `r xx5[[8]]`, coastal:  `r xx5[[11]]`).

Some countries used another approach to set their management objective (lakes: `r xx5[[3]]`, rivers:  `r xx5[[6]]`, transitional:  `r xx5[[9]]`, coastal:  `r xx5[[12]]`, marine: `r xx5[[14]]`)

Setting nutrient boundaries that are consistent with the G/M boundary will not necessarily equate with achieving good status for BQEs, particularly as G/M boundaries are usually type- rather than site-specific.  Further questions will need to be asked to elucidate how countries address the inevitable mismatches between chemical and ecological status.   If ecology does not achieve good status after nutrient reductions to the G/M boundary, then what actions does each country take?   This could be a situation where “less stringent objectives” are invoked, rather than reviewing evidence specific to a particular water body.   

## Q6. If you have set management objectives for nutrients, please specify which policy the management objectives are meant to address.

```{r Q6, fig.cap="Figure 6. Number of countries responding, overview by water category."}
#select data Q6
PartA.Q6 <- dat.PartA %>% filter(QuestionNo_b==6)

PartA.Q6$ResponseOptionText <- str_trim(PartA.Q6$ResponseOptionText)
  
plotQ6 <- ggplot(PartA.Q6, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotQ6
```

For all water categories except Marine, the WFD is the most important driver, though the Urban Wastewater and Nitrates Directive are also cited by several countries across all water types.   For Marine waters the MFSD and Regional Sea Conventions  are both important.   Those who selected “other” c ited the Industrial Emissions Directive, Common Agricultural Policy and EU regulations on phosphate-free detergents as also influencing management directives, along with national legislation (e.g. SE has a policy on “zero eutrophication”: https://www.sverigesmiljomal.se/environmental-objectives/).    



## Q7. Countries which have not set management objectives for nutrients. 

```{r Q7, fig.cap="Figure 7. Number of countries responding, overview by water category."}
#select data Q6
PartA.Q7 <- dat.PartA %>% filter(QuestionNo_b==7)

PartA.Q7$ResponseOptionText <- str_trim(PartA.Q7$ResponseOptionText)
  
plotQ7 <- ggplot(PartA.Q7, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

#plotQ7

```
```{r Countries detail Q7, include=FALSE}
PartA.Q7long <-pivot_longer(PartA.Q7, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.Q7long <- PartA.Q7long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.count7 <- PartA.Q7long %>%
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

unique(id.count7$WCreply) #levels with country replies
xx7 <-CntryListfunction (id.count7)#countries replying
#xx?[[11]] # use to extract values directly in text
```

Countries which have not set management objectives for nutrients in lakes (`r xx7[[1]]`), rivers, (`r xx7[[2]]`, transitional (`r xx7[[3]]`), coastal (`r xx7[[4]]`), marine (`r xx7[[5]]`).

(see word doc for the reasons and information on how they achieve the WFD and/or MSFD objectives instead for water bodies exposed to nutrient pollution.)
Add here a critical analysis paragraph.

## How are the current nutrient loads estimated? (Part B)
```{r pre-treatment B, include=FALSE}
dat.PartB <- type.convert(dat.PartB) #convert data automatic as appropriate

dat.PartB$BE<-as.integer(dat.PartB$BE)
dat.PartB$BE.Br<-as.integer(dat.PartB$BE.Br)
dat.PartB$BE.Fl<-as.integer(dat.PartB$BE.Fl)

PartB.long <-pivot_longer(dat.PartB, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartB.long <- PartB.long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

PartB.long$Acronym <- as.factor(PartB.long$Acronym)
```

## Q1. Do you calculate current nutrient loads in order to achieve management objectives for nutrients?
```{r BQ1, fig.cap="Figure 7. Number of countries responding."}
PartB.Q1 <- dat.PartB %>% filter(QuestionNo_b==1)

PartB.Q1$ResponseOptionText <- str_trim(PartB.Q1$ResponseOptionText)

plotBQ1 <- ggplot(PartB.Q1, aes(ResponseOptionText, ResponsesPerOption)) +
  scale_x_discrete() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col(aes(fill=ResponseOptionText)) +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotBQ1 + theme(legend.position = "none") 
```
Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q2. If yes, what methods do you use to calculate nutrient loads?
```{r BQ2, fig.cap="Figure 8. Number of countries responding."}
#select data BQ2
PartB.Q2 <- dat.PartB %>% filter(QuestionNo_b==2)

PartB.Q2$ResponseOptionText <- str_trim(PartB.Q2$ResponseOptionText)
  
plotBQ2 <- ggplot(PartB.Q2, aes(ResponseOptionText, ResponsesPerOption)) + 
  scale_x_discrete(limits = rev(c("Loads are calculated from monitoring of nutrient concentrations multiplied with flow in the inflows to the water body", "Loads are calculated from estimates of emissions from different sources", "Both approaches are used", "Other approaches are used")),
                   labels = rev(c("Loads are calculated from monitoring\nof nutrient concentrations multiplied\nwith flow in the inflows to the water body", "Loads are calculated from estimates\nof emissions from different sources", "Both approaches are used", "Other approaches are used"))) +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotBQ2
```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q3. Do you use catchment models that quantify the nutrient emissions/nutrient losses to surface waters?
```{r BQ3, fig.cap="Figure 9. Number of countries responding."}
PartB.Q3 <- dat.PartB %>% filter(QuestionNo_b==3)

PartB.Q3$ResponseOptionText <- str_trim(PartB.Q3$ResponseOptionText)

PartBQ3 <- ggplot(PartB.Q3, aes(ResponseOptionText, ResponsesPerOption)) +
  scale_x_discrete() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col(aes(fill=ResponseOptionText)) +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

PartBQ3 + theme(legend.position = "none") 
```

```{r Countries detail BQ3, include=FALSE}
PartB.Q3long <-pivot_longer(PartB.Q3, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartB.Q3long <- PartB.Q3long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.countBQ3 <- PartB.Q3long %>%
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

unique(id.countBQ3$WCreply) #levels with country replies
xxBQ3 <-CntryListfunction (id.countBQ3)#countries replying
#xx?[[11]] # use [[?]] to extract values directly in text
```

Catchment models that quantify the nutrient emissions/nutrient losses to surface waters are used by 19 countries, namely: `r xxBQ3[[1]]`. (HT-MK) to complete this paragraph extract relevant information from this open question "If yes, please provide a short description of the type of model you use considering the aspects above and provide links to relevant publications/reports where possible)

Add here a critical analysis paragraph.

## Q4. If yes, at what spatial level do the models operate?
```{r BQ4, fig.cap="Figure 10. Number of countries responding."}
#select data BQ2
PartB.Q4 <- dat.PartB %>% filter(QuestionNo_b==5)

PartB.Q4$ResponseOptionText <- str_trim(PartB.Q4$ResponseOptionText)
  
plotBQ4 <- ggplot(PartB.Q4, aes(ResponseOptionText, ResponsesPerOption)) +
  scale_x_discrete(limits = rev(c("Catchment", "Sub-Catchment", "Water Body", "Specific Spatial Unit - see Word Doc for info")),
                   labels = rev(c("Catchment", "Sub-Catchment", "Water Body", "Specific Spatial Unit"))) +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotBQ4

```

Add here a summary paragraph on the graph results. (note regarding the Specific Spatial Unit we need to see Word Doc for additional info)

Add here a critical analysis paragraph.

## Q5. If yes, how large are the uncertainties in these models and how do you deal with these uncertainties?
```{r Countries detail BQ5, include=FALSE}
PartB.Q5 <- dat.PartB %>% filter(QuestionNo_b==6)

PartB.Q5long <-pivot_longer(PartB.Q5, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartB.Q5long <- PartB.Q5long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.countBQ5 <- PartB.Q5long %>%
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

unique(id.countBQ5$WCreply) #levels with country replies
xxBQ5 <-CntryListfunction (id.countBQ5)#countries replying
#xx?[[11]] # use [[?]] to extract values directly in text
```

Uncertainties in the models that are used to quantify the nutrient emissions and/or nutrient losses to surface waters were addressed by 16 of countries, namely: `r xxBQ5[[1]]`. (HT-MK: Complete this summary paragraph on the open answers of countries from the word documents)

Add here a critical analysis paragraph.

## How are measures to reach management objectives for nutrients defined? (Part C)
```{r pre-treatment C, include=FALSE}
dat.PartC <- type.convert(dat.PartC) #convert data automatic as appropriate

dat.PartC$BE<-as.integer(dat.PartC$BE)
dat.PartC$BE.Br<-as.integer(dat.PartC$BE.Br)
dat.PartC$BE.Fl<-as.integer(dat.PartC$BE.Fl)

dat.PartC$WaterCategory<-factor(dat.PartC$WaterCategory, levels=c("Lakes", "Rivers", 
                                         "Transitional", "Coastal", "Marine", 
                                         "All", "Point sources", "Diffuse sources"))#order factor levels

PartC.long <-pivot_longer(dat.PartC, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartC.long <- PartC.long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

PartC.long$Acronym <- as.factor(PartC.long$Acronym)
```


## Q1. When is nutrient reduction decided to be necessary?
(note to self) Attention: marine waters less categorical options but merge if possible
```{r pre-treat CQ1 labels, include=FALSE}
#pretreatment of some answer options: too long  names
#case_when QuestionNo_b ==1 then change the following ResponseOptionText
"When a water body is classified as failing good status according to nutrient-sensitive-biological quality elements (BQEs)" ~ "Water body failing good status according to nutrient-sensitive-BQEs"

"When a water body is classified failing good status according to nutrient good/moderate boundaries (if yes, please specify if this applies to both P and N, or only to the limiting nutrient)" ~"Water body failing good status according to nutrient good/moderate boundaries" 

"When a water body is classified as failing good status according to both nutrient-sensitive-BQEs and nutrients"  ~ "Water body failing good status according to both nutrient-sensitive-BQEs and nutrients"

"When a water body is classified as good status for nutrient-sensitive BQEs, but deteriorating towards moderate due to increasing pressures or impacts (precautionary principle)" ~"Water body in good status for nutrient-sensitive BQEs, but deteriorating towards moderate due to increasing pressures or impacts (precautionary principle)"

"When criterion D5C1 “nutrient concentrations” fails to achieve good status (if yes, please specify if this applies to both P and N, or only to the limiting nutrient)" ~ "When criterion D5C1 “nutrient concentrations” fails to achieve good status"

```

```{r}

```

Add here a summary paragraph on the graph results.
(note to HT and MK: namely considering where option 2 is selected by any MS, check if they have specified if that applied to both P and N, or only to the limiting nutrient, as requested)  
```{r CQ1 option 2 explore, include=FALSE}

#see if here or directly in text
#id countries selecting ResponseOptionRef == "2.2" or "1.4.2" or "1.3.2" or "1.2.2" or "1.1.2"
```

Add here a critical analysis paragraph.

## Q2. Are you undertaking pollution accounting to identify the sources of nutrient pollution?
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q3. If yes, which point and diffuse sources are included?
(HT note to self) here merge into a single graph point and diffuse using factor WaterCategory
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

(open question) Describe the methodology that is used for the pollution accounting.
Add here a summary paragraph on the open answers and add number of countries replying (directly with formula).

Add here a critical analysis paragraph.

## Q4. What criteria and considerations do you use to select which measures to implement in order to reduce nutrient inputs?
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

(from here Yes/No graphs:)
## Q5. Are past policy measures such as a focus on urban waste water treatment plants as opposed to diffuse pollution from agriculture taken into account when planning new measures, e.g. trying to achieve a fair sharing of the nutrient reductions between different sectors according to a pollution account?
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q6. Is the delay in the response of the ecosystem to reduced nutrient inputs taken into account when planning new measures?
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q7. Do you estimate the effect of measures on the reduction of the nutrient load or concentration in water? 
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q8. Do you have mechanisms for managing management objectives for nutrients across political boundaries (either within or between Member States)? 
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q9. Do you consider climate change impacts on nutrient emissions and/or on biological quality elements/MSFD criterion responses to nutrients when planning the nutrient reduction measures?
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q10. Do you consider nature-based solutions (restoring riparian zones, wetlands, flood plains etc.) when planning the nutrient reduction measures? 
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q11. Do you consider technical/biological manipulative solutions (removal of nutrient-rich sediment, oxygenation of bottom water, adding chemicals to the sediment to bind phosphorus, extractive aquaculture using mussels or seaweeds, management fisheries) when planning the nutrient reduction measures? 
```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.
