---
title: "Methods used to set management objectives for nutrients - ECOSTAT Questionnaire"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    highlight: kate
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

ECOSTAT has a mandate to compare methods used by Member States to set management objectives  for nutrients that are compatible with good ecological status and good environmental status.

This questionnaire aims to understand national approaches for setting and implementing management objectives for nutrients and to help Member States and collaborating countries to improve their approaches and, thereby, achieve the objectives of the Water Framework Directive (WFD) and the Marine Strategy Framework Directive (MSFD).

```{r load packages, include=FALSE}
library(tidyverse)
library(tidytext)
library(here)
library(knitr)
library(kableExtra)
```

```{r load data, include=FALSE}
FName.dat <-here::here("Data","PartA.csv")
dat.PartA<- read.csv(file = FName.dat, sep=";", header = TRUE) #fileEncoding="latin1"
```

```{r pre-treatment, include=FALSE}
dat.PartA <- type.convert(dat.PartA) #convert data automatic as appropriate

dat.PartA$BE<-as.integer(dat.PartA$BE)
dat.PartA$BE.Br<-as.integer(dat.PartA$BE.Br)
dat.PartA$BE.Fl<-as.integer(dat.PartA$BE.Fl)

dat.PartA$WaterCategory<-factor(dat.PartA$WaterCategory, levels=c("Lakes","Rivers","Transitional","Coastal","Marine")) #order factor levels

PartA.long <-pivot_longer(dat.PartA, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.long <- PartA.long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

PartA.long$Acronym <- as.factor(PartA.long$Acronym)
```

#### Overview of countries participationg in questionnaire

Twenty-six countries participated in this survey, five of which are land-locked countries.

```{r countries data, }
dat.countries <- read.csv(file = here::here("Data","Countries.csv"), sep=";", header = TRUE)
dat.countries$Type <- factor(dat.countries$Type, levels = c("EU Member States", "EFTA Member States", "EU Candidate Countries"))

#left join countries full name and info
PartA.long <- left_join( PartA.long, dat.countries, by = "Acronym", keep = FALSE) 
PartA.long$Countries <- as.factor(PartA.long$Countries)

Tab1 <- PartA.long %>% filter(Reply == 1)%>%
  select(c(WaterCategory, Acronym, Countries))%>%
  count(WaterCategory, Acronym, Countries)%>%
  mutate(n = ifelse(n < 100, "y", n)) #replace number of replies just to get countries per wat cat

#Table 1  
options(knitr.kable.NA = '')
Tab1 %>% spread(key = WaterCategory, value = n) %>%
  knitr::kable(format = "html",col.names = c("Acronym"," ","Lakes","Rivers","Transitional","Coastal","Marine"), align = "llccccc", caption = "Table 1. Countries* participating in survey per water category.", table.attr = "style='width:70%;'") %>%
  kableExtra::add_header_above(header = c("Country" = 2, "Water Category" = 5),align = "centre")

```

 *Iceland and Norway (EFTA Members) and Turkey (EU candidate) are not EU Member States. 

## How are management objectives for nutrients derived (methods)? (Part A)

## Q1. Have you set management objectives for nutrients for waters bodies that are exposed to nutrient pollution?

```{r Q1,fig.cap="Figure 1. Number of countries responding, overview by water category."}
#select data Q1
PartA.Q1 <- dat.PartA %>% filter(QuestionNo_b==1)

PartA.Q1$ResponseOptionText <- str_trim(PartA.Q1$ResponseOptionText)

#run function to display the sample size at y=0
give.n <- function(x){
  return(c(y = -1, label = x))#length(x)
}
  
plotQ1 <- ggplot(PartA.Q1, aes(reorder_within(ResponseOptionText,ResponseOptionRef,WaterCategory), ResponsesPerOption)) +
  scale_x_reordered() +
  scale_y_continuous(breaks=seq(0, 26, by=2))+ #adjust scale & intervals values
  geom_col() +
  facet_wrap(~ WaterCategory, ncol=1, scales = "free_y") +
  labs (y = expression(paste("No. of countries")),x = NULL) +
  coord_flip()+ 
  theme_classic()+
  theme(strip.background=element_rect(linetype = "blank"))+
  stat_summary(fun.data = give.n, geom = "text", color="darkgray", size=3) #add sample size text

plotQ1
```

```{r Countries detail Q1, include=FALSE}
PartA.Q1long <-pivot_longer(PartA.Q1, 11:39, names_to = "Country", values_to = "Reply") #transpose countries data

PartA.Q1long <- PartA.Q1long %>% mutate(Acronym=Country) %>%
  mutate (Acronym =case_when(
    Acronym == "BE.Br" ~ "BE",
    Acronym == "BE.Fl" ~ "BE",
    Acronym == "BE.Wa" ~ "BE",
    TRUE ~ as.character(Acronym)))

id.count1 <- PartA.Q1long %>% 
  filter(!is.na(Reply)) %>%
  select(c(WaterCategory,ResponseOptionText, Acronym)) %>%
  unite (WCreply, WaterCategory:ResponseOptionText)

#function to extract countries list per reply
CntryListfunction <- function(dat) {
      #i is a unique WaterCategory - ResponseOptionText combination
      #dat is a subset of the data of interest
  WCreply_levels<- unique(dat$WCreply)
  Lists<-list()# prepare a container
       #loop through WaterCategory - ResponseOptionText unique combinations
  for (i in 1:length(WCreply_levels)) {
  acronym_list<- dat$Acronym[which(dat$WCreply==WCreply_levels[i])]
  Lists[i] <- list(acronym_list)
  #print(Lists)
  }
  return(Lists)
}

unique(id.count1$WCreply) #levels with country replies
xx <-CntryListfunction (id.count1)#countries replying
#xx[[11]] # use [[?]] to extract values directly in text
```

The majority of countries have set management objectives for all water bodies exposed to nutrient pollution in all water categories, with few countries considering only some of their nutrient polluted water bodies (`r xx[[2]]` in lakes; `r xx[[5]]` rivers; `r xx[[8]]` in transitional waters; and `r xx[[11]]` in marine waters). Only `r xx[[3]]` in lakes, `r xx[[6]]` in rivers, `r xx[[12]]` in marine waters have not set management objectives for water bodies exposed to nutrient pollution. Results show that France ("FR") does not follow a standardised national approach, varying according to water agency (but confirm this).

(tbc Martyn) Add here a critical analysis paragraph for freshwater categories.
(tbc Heliana) Add here a critical analysis paragraph for marine water categories.
 
## Q2. When setting management objectives for nutrients, is there a focus on reduction of phosphorus, nitrogen, or both or on the limiting nutrient?

```{r Q2, fig.cap="Figure 2. Number of countries responding, overview by water category."}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph. (namely, refer why they did this as they were asked to explicitly mention it)

## Q3. If you have set management objectives for nutrients, please indicate which methodology has been used. 

```{r}

```

Add here a summary paragraph on the graph results.

Countries also provided a short description of the methodology (including links to background reports or publications). According to their description, only XX countries have considered pressure-response relationships: XX, XX,...

Add here a critical analysis paragraph.

## Q4. Do you determine the maximum critical load or a load reduction target that is compatible with good status for nutrient sensitive BQEs? 

```{r}

```

Add here a summary paragraph on the graph results.

Add here a summary paragraph on (What methodology do you use?) open replies.

Add here a critical analysis paragraph of the graph and of the open question.

## Q5. Is there a relationship between the management objectives set for nutrients and the nutrient boundaries set for ecological status? 

```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q6. If you have set management objectives for nutrients, please specify which policy the management objectives are meant to address.

```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.

## Q7. Countries which have not set management objectives for nutrients. 

```{r}

```

Add here a summary paragraph on the graph results.

(see word doc for the reasons and information on how they achieve the WFD and/or MSFD objectives instead for water bodies exposed to nutrient pollution.)
Add here a critical analysis paragraph.

## BLA BLA BLA (Part B) 


```{r}

```

Add here a summary paragraph on the graph results.

Add here a critical analysis paragraph.
